---
title: "p8105_hw1_yl5828"
author: "Leah Li"
date: "2025-09-19"
output: html_document
---

# HW1

#### Problem 1

We first install the required "moderndive" package in Console.

```{r install package and dataset}
# Load the dataset
library(moderndive)
library(tidyverse)
data("early_january_weather")
```

```{r check dataset}
help("early_january_weather")
```

The dataset `early_january_weather` from the `moderndive` package has character variable **origin**, integer variables **year, month, day, hour**, double class variables **temp, dewp, humid, wind_dir, wind_speed, wind_gust, precip, precip, visib**, and date and time variable **time_hour**. 

The dataset has **`r nrow(early_january_weather) `** rows and **`r ncol (early_january_weather)`** columns. 

The mean temperature is **`r mean(early_january_weather$temp)`** F.


Now, we would like to generate a scatterplot of temperature versus time.

```{r scatterplot}
weather_plot <- ggplot(
  early_january_weather,
  aes(x = time_hour, y = temp, color = humid) ) + 
  geom_point(alpha = .5) +
  theme(legend.position = "bottom")

# Display the plot
print(weather_plot)
```
Several patterns are apparent in this plot:

1.  **Diurnal Cycle:** There is a clear and repeating cyclical pattern in the temperature. The temperature consistently rises during the daytime and falls during the nighttime. This creates the wave-like shape visible across the x-axis, representing the daily temperature cycle.

2.  **Time and Humidity Relationship:** The color of the points reveals a relationship between temperature and time. The humidity increases as the time goes by.

3.  **Time and Temperature Relationship:** While the daily cycle is consistent, the overall peaks and troughs vary from day to day. For instance, the temperatures around January 4th are noticeably lower than those around January 11th, indicating a cold spell followed by a warmer period. Overall, the temperature increases as the time goes by.

Now, we save the scatterplot to our project directory.

```{r extract plot}
ggsave("temp_vs_time_plot.png", plot = weather_plot, width = 10, height = 6, dpi = 300)
```



#### Problem 2

Now, we would like to construct a dataframe with some required vectors.

```{r create dataframe}
# Set a seed for reproducibility
set.seed(123)

# Create the required vectors
normal_sample <- rnorm(10)
logical_vector <- normal_sample > 0
char_vector <- sample(letters, 10)
factor_vector <- factor(sample(c("low", "medium", "high"), 10, replace = TRUE))

set.seed(123)
# Combine them into a data frame
custom_df <- data.frame(
  NormalSample = normal_sample,
  GreaterThan0 = logical_vector,
  Character = char_vector,
  Level = factor_vector
)

set.seed(123)
# Print the data frame
print(custom_df)
```

Now, we want to take the mean of each variable in the dataframe. What works and what doesn’t?

```{r take mean}

```




