---
title: "p8105_hw1_yl5828"
author: "Leah Li"
date: "2025-09-19"
output: html_document
---

# HW1

#### Problem 1

We first install the required "moderndive" package in Console.

```{r install package and dataset}
# Load the dataset
library(moderndive)
library(tidyverse)
data("early_january_weather")
```

```{r check dataset}
help("early_january_weather")
```

The dataset `early_january_weather` from the `moderndive` package has character variable **origin**, integer variables **year, month, day, hour**, double class variables **temp, dewp, humid, wind_dir, wind_speed, wind_gust, precip, precip, visib**, and date and time variable **time_hour**. The dataset has `r nrow(early_january_weather) ` rows and `r ncol (early_january_weather)` columns. The mean temperature is `r mean(early_january_weather$temp)` F.

```{r scatterplot}
weather_plot <- ggplot(
  early_january_weather,
  aes(x = time_hour, y = temp, color = humid) )+ 
  geom_point(alpha = .5) +
  theme(legend.position = "bottom")

# Display the plot in the report
print(weather_plot)
```

Now, we save the scatterplot to our project directory.

```{r extract plot}
ggsave("temp_vs_time_plot.png", plot = weather_plot, width = 10, height = 6, dpi = 300)
```



#### Problem 2








